A thermodynamic view of dusty protoplanetary disks

### 2017-08-03 model.tex
edited in response to AY comments, up to 5.2

Our equilibrium solutions satisfy the hydrostatic constraints of
Eq. \ref{steady_momr}---\ref{steady_momz} but are not in general
steady state solutions to the energy 
equation (\ref{eff_energy}), because the source term $\mathcal{C} \ne0$ for realistic disks.  
Limiting cases with $\mathcal{C} \equiv 0$ include \begin{inparaenum}[1)] 
\item 
  unstratified or 2D, razor-thin disk models with finite dust-gas drag; %(\S\ref{si});   
\item 
  perfectly-coupled dust with $\tstop=0$. %(\S\ref{results}). 
\end{inparaenum} 
The following analyses thus apply strictly to these limiting cases of
$\mathcal{C} = 0$.  Our analysis is also a good approximation when
$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
occur on much longer timescales than the instabilities of interest.

@@

Consider oscillations in the dusty gas with period $T_p$. 
The average rate of work done is 
\begin{align}
  \mathcal{W} = \frac{1}{T_p}\int^{t+T_p}_{t}dt^\prime\int_M P
  \frac{D\upsilon}{Dt^\prime} dm \label{work_def} 
\end{align}
\citep[][Eq. 4.10]{cox67}, 
where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
is the Lagragian derivative, and the 
spatial integral is taken over the total mass $M$ of the mixture and
$dm$ is a mass element.

@@

Now, quite generally, Eq. \ref{lin_mass_full}---\ref{lin_energy_full} 
imply the integrand of the numerator in Eq. \ref{thermal_instability} 
is 
\begin{align} 
  \imag \left(\delta \mathcal{C}
  \nabla\cdot\dd \bm{v}^*\right) = 
  -|\sigma|^2\imag \left(\Delta P 
  \frac{\Delta\rho^*}{\rho}\right). \label{pdv}
\end{align}
%This states that `PdV' work (the right-hand-side) is 
%provided by dust-drag friction, and can be seen explicitly as 
%follows. 
Then combining Eq. \ref{work_real}, \ref{pdv} and
\ref{thermal_instability} 
gives  
\begin{align}
s = \frac{\mathcal{W}}{\mathcal{I}^2}, \label{growth_work}
\end{align}
where we have set $\sigma= - \omega$ in 
Eq. \ref{pdv}. Eq. \ref{growth_work} states that if the average work
done is positive during an oscillation, $\mathcal{W}>0$, then 
then its amplitude would actually grow ($s>0$).  %and not be periodic

>>>>

Our equilibrium solutions satisfy the hydrostatic constraints of
Eq. \ref{steady_momr}---\ref{steady_momz} but are not in general
steady state solutions to the energy 
equation (\ref{eff_energy}), because the source term $\mathcal{C} \ne0$ for realistic disks.  
Limiting cases with $\mathcal{C} \equiv 0$ include \begin{inparaenum}[1)] 
\item 
  unstratified or 2D, razor-thin disk models with finite dust-gas drag; %(\S\ref{si});   
\item 
  perfectly-coupled dust with $\tstop=0$. %(\S\ref{results}). 
\end{inparaenum} 
The following analyses thus apply strictly to these limiting cases of
$\mathcal{C} = 0$ in equilibrium.  Our analysis is also a good approximation when 
$\tstop$ is sufficiently small such that the evolution of the background disk (e.g. dust-settling) 
occur on much longer timescales than the instabilities of interest.

@@

Consider oscillations in the dusty gas with period $T_p$. 
The average rate of work done is 
\begin{align}
  \mathcal{W} = \frac{1}{T_p}\int^{t+T_p}_{t}dt^\prime\int_M P
  \frac{D\upsilon}{Dt^\prime} dm \label{work_def} 
\end{align}
\citep[][see their Eq. 4.10 and related discussions]{cox67}, 
where $\upsilon=1/\rho$ is the specific volume of the mixture, $D/Dt$
is the Lagragian derivative, and the 
spatial integral is taken over the total mass $M$ of the mixture and
$dm$ is a mass element.

@@

Now, quite generally, Eq. \ref{lin_mass_full}---\ref{lin_energy_full} 
imply the integrand of the numerator in Eq. \ref{thermal_instability} 
is 
\begin{align} 
  \imag \left(\delta \mathcal{C}
  \nabla\cdot\dd \bm{v}^*\right) = 
  -|\sigma|^2\imag \left(\Delta P 
  \frac{\Delta\rho^*}{\rho}\right), \label{pdv}
\end{align}
for constant $c_s$. 
%This states that `PdV' work (the right-hand-side) is 
%provided by dust-drag friction, and can be seen explicitly as 
%follows. 
Then combining Eq. \ref{work_real}, \ref{pdv} and
\ref{thermal_instability} 
gives  
\begin{align}
s = \frac{\mathcal{W}}{\mathcal{I}^2}, \label{growth_work}
\end{align}
where we have set $\sigma= - \omega$ in 
Eq. \ref{pdv}. Eq. \ref{growth_work} states that if the average work
done is positive during an oscillation, $\mathcal{W}>0$, then 
then its amplitude would actually grow ($s>0$).  %and not be periodic
